---
title: "WIC ANALYSIS No Code"
subtitle: "An RDD Design"
author: "Colin Heeneman, Emily Frost, Samantha Sepulveda"
date: "`r Sys.Date()`"
format: 
  pdf:
    toc: true
    title-slide: true
editor: visual
---

**Special Supplemental Nutrition Program for Women, Infants, and Children**

PA 528 Dr. Siciliano

\newpage

## Introduction

The motivation for evaluating the Special Supplemental Nutrition Program for Women, Infants, and Children (WIC) stems from its critical role in addressing the nutritional needs of vulnerable populations. Established in 1972, WIC serves low-income mothers, infants, and young children by providing access to healthy food, nutrition education, and healthcare referrals. As of 2024, WIC benefits nearly 53% of all infants born in the United States. Evaluating this program is vital because it addresses issues such as food insecurity, low birth weights, and developmental challenges. Understanding its impact informs policymakers and ensures continued investment in evidence-based solutions to promote public health.

------------------------------------------------------------------------

## Program Overview

WIC was launched in 1972 as a pilot program to combat the nutritional deficiencies observed among low-income families, particularly women and children. It became a permanent federal program in 1975, amid growing societal concerns about equality and public health. The program's primary focus is on early nutritional intervention, which research shows can mitigate medical and developmental issues later in life.

Initially serving 88,000 participants per month, WIC has expanded significantly over the decades, reaching 7.2 million monthly participants by 2000. Its benefits include supplemental foods, nutrition education, and healthcare referrals. The program has shown measurable impacts, such as reducing rates of low birth weights, premature births, and child anemia, as well as improving overall development. Its continued success is attributed to strong congressional backing and data demonstrating its cost-effectiveness in preventing long-term health complications​.

------------------------------------------------------------------------

## Program Theory and Implementation

WIC operates on the premise that addressing nutritional deficiencies among pregnant women and young children leads to improved health outcomes. Grounded in public health research, the program targets key developmental stages with interventions like food vouchers, health screenings, and nutrition education. Studies have shown that these activities promote healthier pregnancies, higher birth weights, and cognitive development. The program also encourages breastfeeding, emphasizing its benefits for immunity and mother-child bonding. These elements reflect a well-articulated theory of change based on established health and nutrition science​.

### Logic Model

**Inputs:** Federal funding, healthcare professionals, and local clinics.\
**Activities:** Nutrition education, food voucher distribution, health screenings, and referrals.\
**Outputs:** Number of participants served, vouchers redeemed, and health screenings conducted.\
**Outcomes:**

-   **Short-term:** Improved nutrition and access to healthcare.

-   **Intermediate:** Healthier pregnancies and reduced child anemia.

-   **Long-term:** Reduced chronic disease prevalence and improved cognitive outcomes.

The logic model visualizes how resources are transformed into tangible activities and outcomes. The linkages between each stage ensure accountability and efficacy​.

```{r logic, echo=FALSE, warning=FALSE, message=FALSE}

 # recreating our logic model using this fun DiagrammeR package
library(DiagrammeR)

grViz("
digraph logic_model {
  graph [layout = dot, rankdir = LR]

  # Node definitions
  node [shape = rectangle, style = filled, fillcolor = grey]
  Inputs [label = 'Inputs']
  Activities [label = 'Activities']
  Outputs [label = 'Outputs']
  Outcomes [label = 'Outcomes']

  node [shape = rectangle, style = filled, fillcolor = pink]
  FederalFunding [label = 'Federal Funding']
  HealthcareProfessionals [label = 'Healthcare Professionals']

  node [shape = rectangle, style = filled, fillcolor = lightblue]
  NutritionEducation [label = 'Nutrition Education']
  FoodVoucherDistribution [label = 'Food Voucher Distribution']
  HealthcareScreenings [label = 'Healthcare Screenings']

  node [shape = rectangle, style = filled, fillcolor = green]
  ParticipantsServed [label = 'Participants Served']
  RedeemedVouchers [label = 'Redeemed Vouchers']
  HealthScreeningsProvided [label = 'Health Screenings Provided']

  node [shape = rectangle, style = filled, fillcolor = gold]
  ShortTerm [label = 'Short Term: Improved Access']
  MediumTerm [label = 'Medium Term: Healthier Pregnancies']
  LongTerm [label = 'Long Term: Reduced Chronic Diseases']

  # Edges between nodes
  Inputs -> Activities
  Activities -> Outputs
  Outputs -> Outcomes

  FederalFunding -> Inputs
  HealthcareProfessionals -> Inputs
  NutritionEducation -> Activities
  FoodVoucherDistribution -> Activities
  HealthcareScreenings -> Activities
  ParticipantsServed -> Outputs
  RedeemedVouchers -> Outputs
  HealthScreeningsProvided -> Outputs
  ShortTerm -> Outcomes
  MediumTerm -> Outcomes
  LongTerm -> Outcomes
}
")


```

## Causation

### Main Outcome

The primary outcome selected is a reduction in low birth weights, as it is directly linked to the program's nutritional interventions and has profound implications for public health​.

### Causal Theory

The causal pathway hypothesizes that WIC participation improves maternal nutrition and healthcare access, influencing birth weight outcomes. WIC's supplemental foods, nutrition education, and healthcare referrals target the main determinants of healthy pregnancies. Improved maternal nutrition reduces risks like low birth weight, while better healthcare access assures timely prenatal interventions.

The model includes socioeconomic status (SES), maternal education, and pre-existing health conditions as control variables to block alternative causal pathways. SES affects access to healthcare and other resources, maternal education influences health literacy and decision-making, and health conditions independently impact birth outcomes. By including these factors, WIC's impact is isolated while confounding is minimized.

The Directed Acyclic Graph (DAG) illustrates relationships with nodes representing variables and arrows indicating causal links. SES, education, and health conditions influence maternal nutrition and healthcare, impacting birth weight directly. The model ensures a robust evaluation of WIC's effects on maternal/child health by controlling for external factor

```{r dag, echo=FALSE, warning=FALSE, message=FALSE}

 # recreating our dag, adding adjustments suggested from Assignment 2 notes

library(dagitty)
library(ggplot2)
library(ggdag)


maternal_dag <- dagify(
  Income ~ Race,
  SubstanceUse ~ Income,
  BirthWeight ~ WIC + MaternalNutrition + HealthcareAccess + Income + SubstanceUse + PEC + Age,
  MaternalNutrition ~ WIC + Income,
  HealthcareAccess ~ WIC + Income + PEC,
  WIC ~ Income,
  exposure = "WIC",
  outcome = "BirthWeight",
  labels = c(
    Race = "Race",
    Income = "Income",
    SubstanceUse = "Substance Use",
    BirthWeight = "Birth Weight",
    WIC = "WIC",
    MaternalNutrition = "Maternal Nutrition",
    HealthcareAccess = "Healthcare Access",
    PEC = "PEC",
    Age = "Age"
  ),
  coords = list(
    x = c(
      Race = 1, Income = 2, SubstanceUse = 3, WIC = 2, MaternalNutrition = 3,
      HealthcareAccess = 4, BirthWeight = 5, PEC = 2, Age = 5
    ),
    y = c(
      Race = 4, Income = 4, SubstanceUse = 3, WIC = 5, MaternalNutrition = 5,
      HealthcareAccess = 3, BirthWeight = 4, PEC = 2, Age = 5
    )
  )
)

# Tidy the DAG and assign node statuses
tidy_dag <- tidy_dagitty(maternal_dag) %>%
  node_status()

# Visualize the DAG
ggdag(tidy_dag, text = FALSE) +
  geom_dag_node(aes(color = status), size = 10) + 
  geom_label(aes(x = x, y = y, label = label), size = 4, label.size = 0.2, fill = "white") + 
  scale_color_manual(
    values = c("exposure" = "pink", "outcome" = "lightblue", "unexposed" = "black")
  ) +
  theme_dag() +
  labs(title = "DAG for Maternal Factors Influencing Birth Weight") 

```

### Hypotheses

Participation in WIC will reduce the incidence of low birth weights. The research question asks: Does WIC participation causally improve birth outcomes among low-income mothers?

------------------------------------------------------------------------

## Data and Methods

### Identification Strategy

Regression Discontinuity (RD) will be performed analyzing participants near the income eligibility threshold, ensuring local comparability and minimizing bias. Threats to validity include manipulation of income data near the threshold and unobserved time-varying factors​

Given the nature of our program, income eligibility is the threshold for cutoff, which allows for causal inference by comparing individuals above and below the threshold. By focusing on a narrow bandwidth of income near the cutoff, RDD ensures local comparability, minimizes selection bias, and enhances internal validity.

This framework accounts for potential selection bias and endogeneity by exposing quasi-random assignment of program eligibility near the threshold. It assumes individuals on either side of the cutoff are similar in all aspects except for program participation, enabling isolation of the causal effects and attributing differences in birth weight outcomes to the WIC involvement. Administrative data will further strengthen the analysis, providing accurate and consistent measures of income, birth weight, and other essential variables, reducing the risk of reporting bias.

Despite its strengths, the study faces threats to internal and external validity, as internal manipulation of income data near the threshold could bias the results if individuals deliberately misreport their income to qualify for WIC. To mitigate this, density plots will be used to check for robustness and irregular income distribution near the cutoff. Also, unobserved time-varying factors may influence outcomes, such as economic conditions or policy changes. These factors will be controlled by including covariates to capture temporal trends.

Externally, RDD's localized nature limits generalizability as its findings primarily apply to individuals near the eligibility threshold. Broader populations with significantly different socioeconomic profiles can experience different effects. Nevertheless, the RDD approach provides robust causal estimates and insights into WICs' efficiency in addressing nutritional deficiencies and improving maternal/infant health outcomes.

### Data

The analysis will use administrative and synthetic data to assess WIC's impact on birthweight. To ensure accuracy, administrative data will be collected from WIC clinics and health departments, including birth weights, maternal nutrition, and SES metrics. While HIPAA laws limit access, partnerships with local facilities ensure feasibility.

Additionally, a synthetic dataset of 1,000 observations will simulate variables such as income, WIC participation, SES, and birth weight. The dataset intends to account for confounders such as maternal education and pre-existing conditions while having mediators such as nutrition. This combined approach ensures a robust method like Regression Discontinuity Design (RDD) can ground its findings in real-world applicability.

------------------------------------------------------------------------

## Synthetic Analysis

### Setup

```{r setup, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}

 # load necessary libraries
library(dagitty)
library(ggdag)
library(ggplot2)
library(dplyr)
library(rdrobust)
library(stargazer)
library(modelsummary)
library(knitr)

```

### Generate Data

```{r load simulated data, echo=FALSE, warning=FALSE, message=FALSE}

simulated_data_optimistic <- read.csv("simulated_wic_data_optimistic.csv")

simulated_data_pessimistic <- read.csv("simulated_wic_data_pessimistic.csv")


```

### Applying RDD

```{r RDD, echo=FALSE, warning=FALSE, message=FALSE}

# Setting the cutoff point
cutoff <- 25

# Optimistic Approach
# Define the running variable (income) and the outcome variable (birth_weight)
running_variable_optimistic <- simulated_data_optimistic$income_optimistic
outcome_variable_optimistic <- simulated_data_optimistic$birth_weight_optimistic

 # ensure sharp design
simulated_data_optimistic$wic_participation_optimistic <- ifelse(simulated_data_optimistic$income_optimistic <= cutoff, 1, 0)

# Perform RDD analysis using rdrobust for the optimistic approach
rdd_results_optimistic <- rdrobust(
  y = outcome_variable_optimistic,
  x = running_variable_optimistic,
  c = cutoff
)

# View the summary of RDD results for the optimistic approach
summary(rdd_results_optimistic)

```

**Visualizing Cutoff For RDD**

```{r VisualizeCutoff, echo=FALSE, warning=FALSE, message=FALSE}
# Prepare the data for visualization using the optimistic approach
visualization_data <- simulated_data_optimistic
visualization_data$wic_participation_factor <- as.factor(visualization_data$wic_participation_optimistic)

# Create the plot with reduced vertical gap between bands
ggplot(visualization_data, aes(x = income_optimistic, y = wic_participation_optimistic, color = wic_participation_factor)) +
  geom_jitter(alpha = 0.5, width = 1, height = 0.05) +  # Add jitter
  geom_vline(xintercept = cutoff, color = "grey", linetype = "dashed") +  # Add cutoff line
  labs(
    x = "Income",
    y = "Probability of Participating in WIC",
    color = "WIC Participation"
  ) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("Not Participating", "Participating"),
    limits = c(0, 1)   
  ) +
  theme_minimal() +
  theme(legend.position = "top") +
  ggtitle("Effect of Income Cutoff on WIC Participation (Reduced Gap)")


```

In this graph demonstrating the cutoff mark for participation (1) and non-participation (0), the dotted line signifies one's eligibility for the WIC program. For individuals with income levels below that line, participation in the program is nearly universal, as shown by the cluster of blue points. For those above the line, participation is nonexistent, demonstrating a sharp discontinuity cutoff for the program based on income levels. The sharp change in participation at the cutoff supports the validity of using RDD to estimate the causal effect of WIC participation on the outcomes of interest (birth weight or premature births) by comparing individuals just below and above the income threshold.

**Visualizing Treatment Effect : Optimistic Model**

```{r VisualizeTE, echo=FALSE, warning=FALSE, message=FALSE}

# Visualizing Treatment Effect for Optimistic Approach

# Create a data frame for visualization
visual_data_optimistic <- data.frame(
  income = running_variable_optimistic,
  birth_weight = outcome_variable_optimistic,
  wic_participation = simulated_data_optimistic$wic_participation_optimistic
)

# Convert birth weight to pounds
visual_data_optimistic$birth_weight_pounds <- visual_data_optimistic$birth_weight / 453.592

# Extract treatment effect estimate from rdrobust
treatment_effect_pounds_optimistic <- (rdd_results_optimistic$Estimate[1] / 453.592)  # Convert effect from grams to pounds

# ggplot2 visualization
ggplot(visual_data_optimistic, aes(x = income, y = birth_weight_pounds, color = factor(wic_participation))) +
  geom_point(alpha = 0.6) +  
  geom_smooth(data = subset(visual_data_optimistic, income <= cutoff), method = "lm", color = "darkblue", se = FALSE) +
  geom_smooth(data = subset(visual_data_optimistic, income > cutoff), method = "lm", color = "darkred", se = FALSE) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +
  labs(
    title = "Regression Discontinuity Plot: Birth Weight vs. Income (Optimistic Approach)",
    x = "Income",
    y = "Birth Weight (pounds)",
    color = "WIC Participation"
  ) +
  annotate(
    "text",
    x = 55,  
    y = max(visual_data_optimistic$birth_weight_pounds) * 0.8,  
    label = paste("Estimated Treatment Effect:", round(treatment_effect_pounds_optimistic, 2), "pounds"),
    size = 4,
    color = "black"
  ) +
  theme_minimal()

```

This graph shows a regression discontinuity plot that explores the relationship between income, WIC participation, and birth weight, where the dotted line represents the income cutoff for program participation, and how participation impacts the birth weight for children/families who do or do not partake in the program. The regression line for participants (blue) and non-participants (red) shows a slight positive slope, indicating that birth weight generally increases with income in both groups. However, the sharp drop at the income cutoff highlights the distinct impact of WIC participation. The estimated treatment effect is labeled as 0.58 pounds, meaning that WIC participation is associated with an increase in birth weight by 0.58 pounds compared to those who do not participate.

### Include Robustness Checks

First, we check for manipulation using density plots. Initial approach makes use of a line graph and then use the histogram approach.

```{r Density Line, echo=FALSE, warning=FALSE, message=FALSE}

# Density Plot (Line Graph Style) for Optimistic Approach

# ggplot2 density plot (line graph style)
ggplot(simulated_data_optimistic, aes(x = income_optimistic)) +
  geom_density(color = "darkblue", fill = "lightblue", alpha = 0.5) +  
  geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +  
  labs(
    title = "Density Plot of Income (Optimistic Approach)",
    x = "Income",
    y = "Frequency"
  ) +
  theme_minimal()


```

This graph utilizes the data that shows the distribution of income for individuals who are eligible for the WIC program and those who are not, using the optimistically created data. The dashed line indicates where the treatment (WIC participation) is determined based on income. The distribution's continuity around this point is crucial for ensuring the validity of the regression discontinuity design. If there is no discontinuity in the density at the threshold, it supports the assumption that income assignment near the threshold is as good as random. The smooth curve near the dashed line suggests that there is no evident manipulation of income to qualify or disqualify for WIC. This strengthens the causal interpretation of the estimated treatment effect from the previous RD analysis.

```{r Density Hist, echo=FALSE, warning=FALSE, message=FALSE}

# Density Plot (Histogram Style) for Optimistic Approach

# Add a variable to indicate whether income is above or below the cutoff
visual_data_optimistic <- simulated_data_optimistic %>%
  mutate(income_side = ifelse(income_optimistic < cutoff, "Treated", "Untreated"))

# ggplot2 histogram
ggplot(visual_data_optimistic, aes(x = income_optimistic, fill = income_side)) +
  geom_histogram(
    binwidth = 2, 
    color = "black", 
    alpha = 0.7, 
    boundary = cutoff  
  ) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +  
  scale_fill_manual(
    values = c("Treated" = "#0F3E67", "Untreated" = "red"),
    name = "Treatment Status"
  ) +
  labs(
    title = "Histogram of Income Around Cutoff (Optimistic Approach)",
    x = "Income",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(legend.position = "top") 


```

Using the optimistically created data, a histogram was created to see the distribution of income between individuals who participate in the WIC program and those who were ineligible to do as their incomes were greater than the cutoff. There is a sharp change in the number of individuals on either side of the cutoff, reflecting the income eligibility criteria for WIC and indicates that households are being clearly sorted into treated and untreated groups based on their income. If the WIC program is effective, we would expect to see significant differences in outcomes (better birth weights or improved nutrition) between the treated and untreated groups, particularly for those just around the income cutoff. The shape of the histogram might also suggest areas for policy adjustments---for instance, if there are many households just above the cutoff, they may benefit from expanded eligibility thresholds.

```{r OptRobustness, echo=FALSE, warning=FALSE, message=FALSE}

# Optimistic Design

# TESTING AT DIFFERENT BANDWIDTHS FOR ROBUSTNESS (Optimistic Approach)

# Bandwidth = 10
rdd_results2_optimistic <- rdrobust(y = outcome_variable_optimistic, x = running_variable_optimistic, c = cutoff, h = 10)

# View the summary of results for bandwidth = 10
summary(rdd_results2_optimistic)

# Bandwidth = 5
rdd_results3_optimistic <- rdrobust(y = outcome_variable_optimistic, x = running_variable_optimistic, c = cutoff, h = 5)

# View the summary of results for bandwidth = 5
summary(rdd_results3_optimistic)



```

***Now for added rigor and fun, we repeat the process with our pessimistic data simulation.***

```{r RDD pess, echo=FALSE, warning=FALSE, message=FALSE}

# Pessimistic Approach
# Define the running variable (income) and the outcome variable (birth_weight)
running_variable_pessimistic <- simulated_data_pessimistic$income_pessimistic
outcome_variable_pessimistic <- simulated_data_pessimistic$birth_weight_pessimistic

# Ensure sharp RDD
simulated_data_pessimistic$wic_participation_pessimistic <- ifelse(simulated_data_pessimistic$income_pessimistic <= cutoff, 1, 0)

# Perform RDD analysis using rdrobust for the pessimistic approach
rdd_results_pessimistic <- rdrobust(
  y = outcome_variable_pessimistic,
  x = running_variable_pessimistic,
  c = cutoff
)

# View the summary of RDD results for the pessimistic approach
summary(rdd_results_pessimistic)


```

```{r Treatment Visualization pess, echo=FALSE, warning=FALSE, message=FALSE}

# Create a data frame for visualization (Pessimistic Approach)
visual_data_pessimistic <- data.frame(
  income = running_variable_pessimistic,
  birth_weight = outcome_variable_pessimistic,
  wic_participation = simulated_data_pessimistic$wic_participation_pessimistic
)

# Convert birth weight to pounds
visual_data_pessimistic$birth_weight_pounds <- visual_data_pessimistic$birth_weight / 453.592

# Extract treatment effect estimate from rdrobust
treatment_effect_pounds_pessimistic <- (rdd_results_pessimistic$Estimate[1] / 453.592)  # Convert effect from grams to pounds

ggplot(visual_data_pessimistic, aes(x = income, y = birth_weight_pounds, color = factor(wic_participation))) +
  geom_point(alpha = 0.6) +
  geom_smooth(data = subset(visual_data_pessimistic, income <= cutoff), method = "lm", color = "#C1C12D", se = FALSE) +
  geom_smooth(data = subset(visual_data_pessimistic, income > cutoff), method = "lm", color = "#7E2DC1", se = FALSE) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c("0" = "#C6A6EE", "1" = "#E5E55D"),  
    name = "WIC Participation"
  ) +
  labs(
    title = "Regression Discontinuity Plot: Birth Weight vs. Income (Pessimistic Approach)",
    x = "Income",
    y = "Birth Weight (pounds)"
  ) +
  annotate(
    "text",
    x = 55, 
    y = max(visual_data_pessimistic$birth_weight_pounds) * 0.8, 
    label = paste("Estimated Treatment Effect:", round(treatment_effect_pounds_pessimistic, 2), "pounds"),
    size = 4,
    color = "black"
  ) +
  theme_minimal()


```

This regression discontinuity plot was created using the pessimistic fake data to analyze the regression discontinuity design for WIC. The estimated treatment effect of WIC participation on birth weight is 0.2 pounds, as indicated. This suggests that, even with the use of the pessimistic data, there is still an increase in birth weight for individuals who partake in the WIC program.

```{r Robust pess, echo=FALSE, warning=FALSE, message=FALSE}

# TESTING AT DIFFERENT BANDWIDTHS FOR ROBUSTNESS (Pessimistic Approach)

# Bandwidth = 10
rdd_results2_pessimistic <- rdrobust(y = outcome_variable_pessimistic, x = running_variable_pessimistic, c = cutoff, h = 10)

# View the summary of results for bandwidth = 10
summary(rdd_results2_pessimistic)

# Bandwidth = 5
rdd_results3_pessimistic <- rdrobust(y = outcome_variable_pessimistic, x = running_variable_pessimistic, c = cutoff, h = 5)

# View the summary of results for bandwidth = 5
summary(rdd_results3_pessimistic)


```

Below is the results of robustness tests for the optimistic and pessimistic approach, showing treatment effect size on birth weight (grams).

```{r TableTE, echo=FALSE, warning=FALSE, message=FALSE}

# Extract treatment effect estimates
treatment_effects <- data.frame(
  Model = c("RDD Optimistic (h=10)", "RDD Optimistic (h=5)", 
            "RDD Pessimistic (h=10)", "RDD Pessimistic (h=5)"),
  Estimate = c(
    rdd_results2_optimistic$Estimate[1],
    rdd_results3_optimistic$Estimate[1],
    rdd_results2_pessimistic$Estimate[1],
    rdd_results3_pessimistic$Estimate[1]
  )
)


kable(treatment_effects, caption = "Treatment Effect Estimates from Different RDD Models (Grams)")

```

------------------------------------------------------------------------

## Conclusion

Our analysis used both optimistic and pessimistic scenarios within a Regression Discontinuity Design (RDD) framework to evaluate the impact of the Special Supplemental Nutrition Program for Women, Infants, and Children (WIC) on birth weight outcomes. Our findings suggest a positive impact of WIC participation on birth weights when interpreted correctly.

### Key Findings:

**Optimistic Scenario:**

-   Treatment Effect: 257.91 grams (95% CI: 131.06 to 384.77)

-   Significant (z = 3.99, p \< 0.001)

Optimistic data simulation revealed a statistically significant positive treatment effect. The positive sign, interpreted after adjusting for RDD calculation conventions, indicates that WIC participation leads to an increase in birth weight.

**Pessimistic Scenario:**

-   Treatment Effect: 87.41 grams (95% CI: -54.30 to 229.13)

-   Not statistically significant (z = 1.21, p = 0.227)

In the pessimistic analysis, the treatment effect was positive but not statistically significant. This suggests that under less favorable assumptions, WIC's effect on birth weight may be modest or uncertain.

### Policy Implications:

These findigs highlight the potential for WIC's positive role in enhancing birth weight outcomes among lower-income mothers, justifying continued program investment and possible expansions.

Future research should include more comprehensive administrative data, explore potential heterogeneity by demographic subgroups, and account for possible measurement errors or unobserved variables influencing birth weight. This would provide a clearer understanding of WIC's impact and inform more targeted policy adjustments.
