---
title: "WIC Synthetic Analysis"
subtitle: "A Regression Discontinuity Evaluation of Birth Weight Outcomes"
author:
  - name: "Colin Heeneman"
  - name: "Emily Frost"
  - name: "Samantha Sepulveda"
date: "December 5, 2024"
date-modified: "April 21, 2026"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "Contents"
    css: wic-styles.css
    embed-resources: true
    smooth-scroll: true
    number-sections: false
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
    fig-width: 8
    fig-height: 5
    fig-align: center
editor: visual
---

```{r global-setup, include=FALSE}
# Global knitr options
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi     = 150,
  out.width = "100%"
)
```

```{r font-setup, include=FALSE}
# Load EB Garamond via showtext for consistent typography in plots.
# Requires the showtext package; falls back gracefully if unavailable.
if (requireNamespace("showtext", quietly = TRUE) &&
    requireNamespace("sysfonts", quietly = TRUE)) {
  sysfonts::font_add_google("EB Garamond", "eb_garamond")
  showtext::showtext_auto()
  showtext::showtext_opts(dpi = 150)
  plot_font <- "eb_garamond"
} else {
  plot_font <- "serif"
}
```

```{r palette, include=FALSE}
# Forest color palette — used consistently across all figures
p_charcoal  <- "#333d29"  # darkest; headings, borders
p_ebony     <- "#414833"  # dark green; main fills
p_olive     <- "#656d4a"  # mid green; secondary elements
p_sage2     <- "#a4ac86"  # light sage; fills, highlights
p_sage      <- "#c2c5aa"  # lightest sage; backgrounds
p_beige     <- "#b6ad90"  # warm neutral; grid, minor elements
p_saddle    <- "#7f4f24"  # warm brown; accent / untreated
p_toffee    <- "#936639"  # toffee; secondary accent
p_bg        <- "#f7f5f0"  # off-white page background
```

------------------------------------------------------------------------

**Special Supplemental Nutrition Program for Women, Infants, and Children** \| PA 528 --- Dr. Siciliano

------------------------------------------------------------------------

## Introduction

The motivation for evaluating the Special Supplemental Nutrition Program for Women, Infants, and Children (WIC) stems from its critical role in addressing the nutritional needs of vulnerable populations. Established in 1972, WIC serves low-income mothers, infants, and young children by providing access to healthy food, nutrition education, and healthcare referrals. As of 2024, WIC benefits nearly 53% of all infants born in the United States. Evaluating this program is vital because it addresses issues such as food insecurity, low birth weights, and developmental challenges. Understanding its impact informs policymakers and ensures continued investment in evidence-based solutions to promote public health.

------------------------------------------------------------------------

## Program Overview

WIC was launched in 1972 as a pilot program to combat the nutritional deficiencies observed among low-income families, particularly women and children. It became a permanent federal program in 1975, amid growing societal concerns about equity and public health. The program's primary focus is on early nutritional intervention, which research demonstrates can mitigate medical and developmental issues later in life.

Initially serving 88,000 participants per month, WIC expanded significantly over the decades, reaching 7.2 million monthly participants by 2000. Its benefits include supplemental foods, nutrition education, and healthcare referrals. The program has shown measurable impacts, including reduced rates of low birth weights, premature births, and child anemia, as well as improved overall development. Its continued success is attributed to strong congressional backing and data demonstrating its cost-effectiveness in preventing long-term health complications.

------------------------------------------------------------------------

## Program Theory and Implementation

WIC operates on the premise that addressing nutritional deficiencies among pregnant women and young children leads to improved health outcomes. Grounded in public health research, the program targets key developmental stages with interventions such as food vouchers, health screenings, and nutrition education. Studies have shown that these activities promote healthier pregnancies, higher birth weights, and cognitive development. The program also encourages breastfeeding, emphasizing its benefits for immunity and mother-child bonding. These elements reflect a well-articulated theory of change rooted in established health and nutrition science.

### Logic Model

**Inputs:** Federal funding, healthcare professionals, and local clinics. **Activities:** Nutrition education, food voucher distribution, health screenings, and referrals. **Outputs:** Number of participants served, vouchers redeemed, and health screenings conducted. **Outcomes:**

-   **Short-term:** Improved nutrition and access to healthcare.
-   **Intermediate:** Healthier pregnancies and reduced child anemia.
-   **Long-term:** Reduced chronic disease prevalence and improved cognitive outcomes.

The logic model below visualizes how resources are transformed into tangible activities and outcomes. The linkages between each stage ensure accountability and program efficacy.

```{r logic-model, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.2}
# Static ggplot2 logic model — no JavaScript dependency
library(ggplot2)

bw <- 1.00   # box half-width  (data units)
bh <- 0.32   # item box half-height
hh <- 0.40   # header box half-height

cx <- c(2.0, 5.5, 9.0, 12.5)  # column x-centers

# ── Headers ───────────────────────────────────────────────────────────────
hdr <- data.frame(
  x     = cx,
  y     = 5.8,
  label = c("Inputs", "Activities", "Outputs", "Outcomes"),
  fill  = c(p_charcoal, p_ebony, p_olive, p_saddle),
  stringsAsFactors = FALSE
)

# ── Item boxes ────────────────────────────────────────────────────────────
items <- data.frame(
  x     = c(
    rep(cx[1], 2),
    rep(cx[2], 3),
    rep(cx[3], 3),
    rep(cx[4], 3)
  ),
  y     = c(
    4.85, 3.95,
    5.05, 4.15, 3.25,
    5.05, 4.15, 3.25,
    5.05, 4.15, 3.25
  ),
  label = c(
    "Federal Funding",        "Healthcare\nProfessionals",
    "Nutrition\nEducation",   "Food Voucher\nDistribution", "Healthcare\nScreenings",
    "Participants\nServed",   "Redeemed\nVouchers",         "Health Screenings\nProvided",
    "Short Term:\nImproved Access",
    "Medium Term:\nHealthier Pregnancies",
    "Long Term:\nReduced Chronic\nDiseases"
  ),
  fill  = c(
    p_ebony,  p_ebony,
    p_olive,  p_olive,  p_olive,
    p_sage2,  p_sage2,  p_sage2,
    p_beige,  p_toffee, p_saddle
  ),
  tcol  = c(
    "#f7f5f0", "#f7f5f0",
    "#f7f5f0", "#f7f5f0", "#f7f5f0",
    "#333d29", "#333d29", "#333d29",
    "#333d29", "#f7f5f0", "#f7f5f0"
  ),
  stringsAsFactors = FALSE
)

# ── Column-to-column connector arrows ─────────────────────────────────────
col_midy <- c(
  mean(c(4.85, 3.95)),
  mean(c(5.05, 4.15, 3.25)),
  mean(c(5.05, 4.15, 3.25)),
  mean(c(5.05, 4.15, 3.25))
)

arrows_df <- data.frame(
  x    = cx[1:3] + bw + 0.18,
  xend = cx[2:4] - bw - 0.18,
  y    = col_midy[1:3],
  yend = col_midy[2:4]
)

# ── Plot ──────────────────────────────────────────────────────────────────
ggplot() +
  geom_rect(
    data = items,
    aes(xmin = x - bw, xmax = x + bw,
        ymin = y - bh,  ymax = y + bh,
        fill = fill),
    color = p_bg, linewidth = 0.35
  ) +
  geom_rect(
    data = hdr,
    aes(xmin = x - bw, xmax = x + bw,
        ymin = y - hh,  ymax = y + hh,
        fill = fill),
    color = p_bg, linewidth = 0.5
  ) +
  scale_fill_identity() +
  geom_text(
    data = items,
    aes(x = x, y = y, label = label, color = tcol),
    size = 2.7, family = plot_font, lineheight = 0.88
  ) +
  geom_text(
    data = hdr,
    aes(x = x, y = y, label = label),
    color = "#f7f5f0", size = 3.3, fontface = "bold", family = plot_font
  ) +
  scale_color_identity() +
  geom_segment(
    data = arrows_df,
    aes(x = x, xend = xend, y = y, yend = yend),
    color = p_beige, linewidth = 0.7,
    arrow = arrow(length = unit(0.18, "cm"), type = "closed")
  ) +
  coord_cartesian(xlim = c(0.8, 13.7), ylim = c(2.75, 6.4)) +
  theme_void(base_family = plot_font) +
  theme(
    plot.background = element_rect(fill = p_bg, color = NA),
    plot.margin     = margin(14, 14, 14, 14)
  )
```

------------------------------------------------------------------------

## Causal Framework

### Primary Outcome

The primary outcome of interest is a reduction in low birth weights. This outcome is directly linked to WIC's nutritional interventions and carries profound implications for public health, including downstream effects on infant mortality, developmental trajectories, and long-term healthcare costs.

### Causal Theory

The causal pathway hypothesizes that WIC participation improves maternal nutrition and healthcare access, which in turn influences birth weight outcomes. WIC's supplemental foods, nutrition education, and healthcare referrals target the principal determinants of healthy pregnancies. Improved maternal nutrition reduces risks associated with low birth weight, while enhanced healthcare access facilitates timely prenatal interventions.

The model incorporates socioeconomic status (SES), maternal education, and pre-existing health conditions as control variables to block alternative causal pathways. SES affects access to healthcare and other resources; maternal education influences health literacy and decision-making; and pre-existing health conditions independently impact birth outcomes. By conditioning on these factors, the analysis isolates WIC's effect while minimizing confounding.

The Directed Acyclic Graph (DAG) below illustrates these relationships, with nodes representing variables and arrows indicating causal linkages.

```{r dag, echo=FALSE, warning=FALSE, message=FALSE}
library(dagitty)
library(ggplot2)
library(ggdag)

maternal_dag <- dagify(
  Income ~ Race,
  SubstanceUse ~ Income,
  BirthWeight ~ WIC + MaternalNutrition + HealthcareAccess + Income + SubstanceUse + PEC + Age,
  MaternalNutrition ~ WIC + Income,
  HealthcareAccess ~ WIC + Income + PEC,
  WIC ~ Income,
  exposure = "WIC",
  outcome = "BirthWeight",
  labels = c(
    Race             = "Race",
    Income           = "Income",
    SubstanceUse     = "Substance Use",
    BirthWeight      = "Birth Weight",
    WIC              = "WIC",
    MaternalNutrition = "Maternal Nutrition",
    HealthcareAccess  = "Healthcare Access",
    PEC              = "PEC",
    Age              = "Age"
  ),
  coords = list(
    x = c(
      Race = 1, Income = 2, SubstanceUse = 3, WIC = 2, MaternalNutrition = 3,
      HealthcareAccess = 4, BirthWeight = 5, PEC = 2, Age = 5
    ),
    y = c(
      Race = 4, Income = 4, SubstanceUse = 3, WIC = 5, MaternalNutrition = 5,
      HealthcareAccess = 3, BirthWeight = 4, PEC = 2, Age = 5
    )
  )
)

tidy_dag <- tidy_dagitty(maternal_dag) %>%
  node_status()

ggdag(tidy_dag, text = FALSE) +
  geom_dag_edges(edge_color = p_beige) +
  geom_dag_node(aes(color = status), size = 10) +
  geom_label(
    aes(x = x, y = y, label = label),
    size       = 3.8,
    label.size = 0.2,
    fill       = p_bg,
    color      = p_charcoal,
    family     = plot_font
  ) +
  scale_color_manual(
    values = c(
      "exposure"  = p_olive,
      "outcome"   = p_saddle,
      "latent"    = p_ebony
    ),
    na.value = p_ebony,
    labels   = c("exposure" = "Exposure (WIC)", "outcome" = "Outcome (Birth Weight)", "latent" = "Covariate"),
    name     = NULL
  ) +
  theme_dag(base_family = plot_font) +
  theme(
    plot.background  = element_rect(fill = p_bg,  color = NA),
    panel.background = element_rect(fill = p_bg,  color = NA),
    legend.background = element_rect(fill = p_bg, color = NA),
    legend.text  = element_text(color = p_charcoal, size = 10, family = plot_font),
    legend.position = "bottom"
  ) +
  labs(title = "DAG: Maternal Factors Influencing Birth Weight")
```

### Research Hypothesis

**Does WIC participation causally improve birth outcomes among low-income mothers?**

The specific hypothesis is that participation in WIC reduces the incidence of low birth weights, holding constant socioeconomic and health-related confounders identified in the DAG above.

------------------------------------------------------------------------

## Data and Methods

### Identification Strategy

Regression Discontinuity Design (RDD) is used to estimate the causal effect of WIC participation on birth weight. Because WIC eligibility is determined by an income threshold, individuals just below and just above this threshold are treated as locally comparable, enabling credible causal inference without randomization.

The income eligibility threshold serves as the cutoff, creating a sharp discontinuity in program participation. By focusing on a narrow bandwidth of income near the cutoff, RDD ensures local comparability, minimizes selection bias, and enhances internal validity. The design assumes that individuals on either side of the cutoff are similar in all pre-treatment characteristics except for program participation, attributing observed differences in birth weight outcomes to WIC involvement.

Two potential threats to validity are addressed. First, manipulation of income data near the threshold could bias results if individuals strategically misreport income to qualify. Density plots are used to test for bunching at the cutoff. Second, unobserved time-varying factors, such as economic conditions or concurrent policy changes, may confound outcomes; covariates are included to capture relevant temporal trends.

While the RDD design provides strong internal validity, its localized nature limits generalizability beyond the population near the income threshold. Populations with substantially different socioeconomic profiles may experience different program effects.

### Data

The analysis draws on a synthetic dataset of 1,000 observations designed to mirror the structure of administrative WIC records. The dataset includes simulated measures of income (the running variable), WIC participation status, birth weight (the primary outcome), maternal race, pre-existing health conditions, maternal nutrition, healthcare access, and maternal age. Two scenarios are analyzed in parallel: an **optimistic** simulation reflecting stronger program effects and a **pessimistic** simulation with more conservative assumptions. This dual-scenario approach allows assessment of how sensitive the estimated treatment effect is to underlying data-generating assumptions.

```{r setup-libraries, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}
library(dagitty)
library(ggdag)
library(ggplot2)
library(dplyr)
library(rdrobust)
library(knitr)
```

```{r load-data, echo=FALSE, warning=FALSE, message=FALSE}
simulated_data_optimistic  <- read.csv("simulated_wic_data_optimistic.csv")
simulated_data_pessimistic <- read.csv("simulated_wic_data_pessimistic.csv")
```

------------------------------------------------------------------------

## Synthetic Analysis

### Shared Plot Theme

```{r theme-wic, echo=FALSE}
# Custom ggplot2 theme applied consistently across all figures
theme_wic <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = plot_font) +
  theme(
    plot.background   = element_rect(fill = p_bg,  color = NA),
    panel.background  = element_rect(fill = p_bg,  color = NA),
    panel.grid.major  = element_line(color = p_sage, linewidth = 0.35),
    panel.grid.minor  = element_line(color = p_sage, linewidth = 0.15),
    plot.title        = element_text(size = 13, face = "bold",   color = p_charcoal, margin = margin(b = 6)),
    plot.subtitle     = element_text(size = 10, face = "italic", color = p_olive,    margin = margin(b = 8)),
    axis.title        = element_text(size = 10, color = p_ebony),
    axis.text         = element_text(size = 9,  color = p_olive),
    legend.background = element_rect(fill = p_bg, color = NA),
    legend.key        = element_rect(fill = p_bg, color = NA),
    legend.text       = element_text(size = 9, color = p_charcoal),
    legend.title      = element_text(size = 9, color = p_charcoal, face = "bold"),
    plot.caption      = element_text(size = 8, color = p_beige, hjust = 0, margin = margin(t = 8))
  )
}
```

### RDD --- Optimistic Scenario

```{r rdd-optimistic, echo=FALSE, warning=FALSE, message=FALSE}
cutoff <- 25

running_variable_optimistic <- simulated_data_optimistic$income_optimistic
outcome_variable_optimistic  <- simulated_data_optimistic$birth_weight_optimistic

simulated_data_optimistic$wic_participation_optimistic <-
  ifelse(simulated_data_optimistic$income_optimistic <= cutoff, 1, 0)

rdd_results_optimistic <- rdrobust(
  y = outcome_variable_optimistic,
  x = running_variable_optimistic,
  c = cutoff
)

summary(rdd_results_optimistic)
```

#### Income Cutoff and Participation

```{r viz-cutoff, echo=FALSE, warning=FALSE, message=FALSE}
visualization_data <- simulated_data_optimistic
visualization_data$wic_participation_factor <-
  factor(visualization_data$wic_participation_optimistic,
         levels = c(0, 1), labels = c("Not Participating", "Participating"))

ggplot(visualization_data,
       aes(x = income_optimistic, y = wic_participation_optimistic,
           color = wic_participation_factor)) +
  geom_jitter(alpha = 0.45, width = 0.8, height = 0.04, size = 1.5) +
  geom_vline(xintercept = cutoff, color = p_toffee, linetype = "dashed", linewidth = 0.8) +
  scale_color_manual(
    values = c("Not Participating" = p_saddle, "Participating" = p_ebony),
    name   = "WIC Status"
  ) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("Not Participating", "Participating"),
    limits = c(-0.05, 1.05)
  ) +
  labs(
    title    = "Sharp Discontinuity at the WIC Income Threshold",
    subtitle = "Individuals below the cutoff are universally enrolled; those above are ineligible",
    x        = "Income",
    y        = NULL,
    caption  = "Dashed line indicates the income eligibility cutoff (25)."
  ) +
  theme_wic() +
  theme(legend.position = "top")
```

The sharp change in participation at the income cutoff is consistent with a well-implemented sharp RDD. Individuals below the threshold participate nearly universally, while those above do not participate at all. This clean discontinuity is a prerequisite for using RDD to estimate causal effects.

#### Treatment Effect Visualization

```{r viz-te-optimistic, echo=FALSE, warning=FALSE, message=FALSE}
visual_data_optimistic <- data.frame(
  income           = running_variable_optimistic,
  birth_weight     = outcome_variable_optimistic,
  wic_participation = simulated_data_optimistic$wic_participation_optimistic
)
visual_data_optimistic$birth_weight_pounds    <- visual_data_optimistic$birth_weight / 453.592
treatment_effect_pounds_optimistic <- rdd_results_optimistic$Estimate[1] / 453.592

ggplot(visual_data_optimistic,
       aes(x = income, y = birth_weight_pounds,
           color = factor(wic_participation))) +
  geom_point(alpha = 0.45, size = 1.4) +
  geom_smooth(
    data   = subset(visual_data_optimistic, income <= cutoff),
    method = "lm", color = p_ebony, se = FALSE, linewidth = 1
  ) +
  geom_smooth(
    data   = subset(visual_data_optimistic, income > cutoff),
    method = "lm", color = p_saddle, se = FALSE, linewidth = 1
  ) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = p_toffee, linewidth = 0.8) +
  scale_color_manual(
    values = c("0" = p_toffee, "1" = p_sage2),
    labels = c("0" = "Not Participating", "1" = "Participating"),
    name   = "WIC Status"
  ) +
  annotate(
    "text",
    x = 57, y = max(visual_data_optimistic$birth_weight_pounds, na.rm = TRUE) * 0.82,
    label  = paste0("Estimated effect: +", round(treatment_effect_pounds_optimistic, 2), " lbs"),
    size   = 3.6, color = p_charcoal, fontface = "italic", family = plot_font
  ) +
  labs(
    title    = "Regression Discontinuity: Birth Weight vs. Income — Optimistic Scenario",
    subtitle = "Slope fitted separately on each side of the income threshold",
    x        = "Income",
    y        = "Birth Weight (pounds)",
    caption  = "Vertical dashed line denotes the eligibility cutoff. Regression lines fitted by OLS within each side."
  ) +
  theme_wic() +
  theme(legend.position = "top")
```

The regression lines estimated separately on each side of the income threshold reveal a visible jump in predicted birth weight at the cutoff. The estimated treatment effect of **`r round(treatment_effect_pounds_optimistic, 2)` pounds** reflects the local average treatment effect for individuals near the eligibility threshold.

### Robustness Checks --- Manipulation Testing

A central assumption of RDD is that individuals cannot precisely manipulate the running variable to sort themselves just below the threshold. Density plots near the cutoff provide a visual diagnostic.

```{r density-line, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(simulated_data_optimistic, aes(x = income_optimistic)) +
  geom_density(fill = p_sage2, color = p_ebony, alpha = 0.6, linewidth = 0.8) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = p_toffee, linewidth = 0.8) +
  labs(
    title    = "Income Density — Optimistic Scenario",
    subtitle = "Smooth density near the cutoff supports the no-manipulation assumption",
    x        = "Income",
    y        = "Density",
    caption  = "Dashed line = eligibility cutoff."
  ) +
  theme_wic()
```

```{r density-hist, echo=FALSE, warning=FALSE, message=FALSE}
visual_data_optimistic_hist <- simulated_data_optimistic %>%
  mutate(income_side = ifelse(income_optimistic < cutoff, "Eligible (Treated)", "Ineligible (Untreated)"))

ggplot(visual_data_optimistic_hist, aes(x = income_optimistic, fill = income_side)) +
  geom_histogram(
    binwidth  = 2,
    color     = p_bg,
    alpha     = 0.85,
    boundary  = cutoff
  ) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = p_charcoal, linewidth = 0.8) +
  scale_fill_manual(
    values = c("Eligible (Treated)" = p_ebony, "Ineligible (Untreated)" = p_saddle),
    name   = "Treatment Status"
  ) +
  labs(
    title    = "Income Distribution Around Cutoff — Optimistic Scenario",
    subtitle = "No bunching immediately below the threshold; supports validity of RDD design",
    x        = "Income",
    y        = "Count",
    caption  = "Bars shaded by treatment status."
  ) +
  theme_wic() +
  theme(legend.position = "top")
```

Both the density curve and the histogram show a smooth, continuous distribution of income near the cutoff, with no evidence of bunching just below the threshold. This supports the assumption that individuals are not manipulating their income to gain program eligibility.

### Bandwidth Sensitivity --- Optimistic Scenario

```{r bandwidth-optimistic, echo=FALSE, warning=FALSE, message=FALSE}
rdd_results2_optimistic <- rdrobust(
  y = outcome_variable_optimistic, x = running_variable_optimistic, c = cutoff, h = 10
)
summary(rdd_results2_optimistic)

rdd_results3_optimistic <- rdrobust(
  y = outcome_variable_optimistic, x = running_variable_optimistic, c = cutoff, h = 5
)
summary(rdd_results3_optimistic)
```

### RDD --- Pessimistic Scenario

The pessimistic simulation uses more conservative assumptions about WIC's causal effect, providing a lower bound on the plausible range of treatment effects.

```{r rdd-pessimistic, echo=FALSE, warning=FALSE, message=FALSE}
running_variable_pessimistic <- simulated_data_pessimistic$income_pessimistic
outcome_variable_pessimistic  <- simulated_data_pessimistic$birth_weight_pessimistic

simulated_data_pessimistic$wic_participation_pessimistic <-
  ifelse(simulated_data_pessimistic$income_pessimistic <= cutoff, 1, 0)

rdd_results_pessimistic <- rdrobust(
  y = outcome_variable_pessimistic,
  x = running_variable_pessimistic,
  c = cutoff
)

summary(rdd_results_pessimistic)
```

```{r viz-te-pessimistic, echo=FALSE, warning=FALSE, message=FALSE}
visual_data_pessimistic <- data.frame(
  income            = running_variable_pessimistic,
  birth_weight      = outcome_variable_pessimistic,
  wic_participation = simulated_data_pessimistic$wic_participation_pessimistic
)
visual_data_pessimistic$birth_weight_pounds     <- visual_data_pessimistic$birth_weight / 453.592
treatment_effect_pounds_pessimistic <- rdd_results_pessimistic$Estimate[1] / 453.592

ggplot(visual_data_pessimistic,
       aes(x = income, y = birth_weight_pounds,
           color = factor(wic_participation))) +
  geom_point(alpha = 0.45, size = 1.4) +
  geom_smooth(
    data   = subset(visual_data_pessimistic, income <= cutoff),
    method = "lm", color = p_ebony, se = FALSE, linewidth = 1
  ) +
  geom_smooth(
    data   = subset(visual_data_pessimistic, income > cutoff),
    method = "lm", color = p_saddle, se = FALSE, linewidth = 1
  ) +
  geom_vline(xintercept = cutoff, linetype = "dashed", color = p_toffee, linewidth = 0.8) +
  scale_color_manual(
    values = c("0" = p_toffee, "1" = p_sage2),
    labels = c("0" = "Not Participating", "1" = "Participating"),
    name   = "WIC Status"
  ) +
  annotate(
    "text",
    x = 57, y = max(visual_data_pessimistic$birth_weight_pounds, na.rm = TRUE) * 0.82,
    label  = paste0("Estimated effect: +", round(treatment_effect_pounds_pessimistic, 2), " lbs"),
    size   = 3.6, color = p_charcoal, fontface = "italic", family = plot_font
  ) +
  labs(
    title    = "Regression Discontinuity: Birth Weight vs. Income — Pessimistic Scenario",
    subtitle = "More conservative assumptions yield a smaller, uncertain treatment effect",
    x        = "Income",
    y        = "Birth Weight (pounds)",
    caption  = "Vertical dashed line denotes the eligibility cutoff. Regression lines fitted by OLS within each side."
  ) +
  theme_wic() +
  theme(legend.position = "top")
```

Under the pessimistic data-generating process, the estimated treatment effect is **`r round(treatment_effect_pounds_pessimistic, 2)` pounds** --- positive in direction but considerably smaller than the optimistic estimate.

### Bandwidth Sensitivity --- Pessimistic Scenario

```{r bandwidth-pessimistic, echo=FALSE, warning=FALSE, message=FALSE}
rdd_results2_pessimistic <- rdrobust(
  y = outcome_variable_pessimistic, x = running_variable_pessimistic, c = cutoff, h = 10
)
summary(rdd_results2_pessimistic)

rdd_results3_pessimistic <- rdrobust(
  y = outcome_variable_pessimistic, x = running_variable_pessimistic, c = cutoff, h = 5
)
summary(rdd_results3_pessimistic)
```

### Treatment Effect Summary Table

The table below consolidates robustness test results across both scenarios and bandwidth specifications, reporting treatment effect estimates in grams.

```{r table-te, echo=FALSE, warning=FALSE, message=FALSE}
treatment_effects <- data.frame(
  Model    = c(
    "Optimistic — h = 10",
    "Optimistic — h = 5",
    "Pessimistic — h = 10",
    "Pessimistic — h = 5"
  ),
  Estimate_g = round(c(
    rdd_results2_optimistic$Estimate[1],
    rdd_results3_optimistic$Estimate[1],
    rdd_results2_pessimistic$Estimate[1],
    rdd_results3_pessimistic$Estimate[1]
  ), 2)
)

colnames(treatment_effects) <- c("Model", "Estimate (grams)")

knitr::kable(
  treatment_effects,
  caption = "Table 1. Treatment effect estimates from bandwidth sensitivity analysis.",
  align   = c("l", "r")
)
```

------------------------------------------------------------------------

## Conclusion

This analysis applied a Regression Discontinuity Design across both optimistic and pessimistic simulated scenarios to evaluate the effect of WIC participation on infant birth weight among mothers near the income eligibility threshold.

### Key Findings

**Optimistic Scenario:** The estimated treatment effect is 257.91 grams (95% CI: 131.06 to 384.77), and is statistically significant (z = 3.99, p \< 0.001). Under optimistic data-generating assumptions, WIC participation is associated with a meaningful and reliably estimated increase in birth weight.

**Pessimistic Scenario:** The estimated treatment effect is 87.41 grams (95% CI: −54.30 to 229.13), and is not statistically significant (z = 1.21, p = 0.227). Under more conservative assumptions, the effect is positive in direction but uncertain, suggesting the program's impact may be modest or dependent on implementation quality and participant compliance.

The bandwidth sensitivity analysis produces consistent directional findings across both scenarios, lending additional credibility to the RDD design. The density diagnostics show no evidence of income manipulation near the threshold, supporting the identifying assumptions of the design.

### Policy Implications

These findings highlight WIC's potential to improve birth weight outcomes for low-income mothers and justify continued program investment. Even in the pessimistic case, the estimated effect is positive and clinically plausible. Future evaluations should incorporate administrative rather than synthetic data, explore potential heterogeneity by race, maternal age, and geography, and account for possible measurement error in income reporting. Expanding program eligibility thresholds modestly upward may also extend these benefits to households currently excluded but demographically similar to current participants.
